Estimating duration using survival analysis

### 2023-10-04 cis-perfect-testing.tex
add RW2 prior predictive

Smoothing priors encode the biological consideration that the hazard should not vary significantly from day-to-day.
It is the Bayesian equivalent of a penalised likelihood, used frequently in the non-parametric frequentist setting~\autocite[e.g.:][]{bacchettiNonparametric}.
Many possible forms of smoothing priors exist, including splines, Gaussian Processes, and random walks.
I use a second-order random walk because it is simple and produces plausible prior predictive distributions.

A second-order random walk prior encodes that the hazard should be linearly changing with some random changes at each time step.
Specifically, the difference on the logit scale between $\lambda_t$ to $\lambda_{t+1}$ is the difference between $\lambda_{t-1}$ and $\lambda_t$ plus some Gaussian-distributed noise.
Formally:
\begin{align}
  \logit\lambda_{t+1}
  &= \logit\lambda_t + (\logit\lambda_t - \logit\lambda_{t-1}) + \sigma \epsilon_t &\text{for $t \geq 2$} \\
  &= 2\logit\lambda_t - \logit\lambda_{t-1} + \sigma \epsilon_t \\
  \epsilon_t &\dist N(0, 1) &\text{for $t \geq 2$}  \\
  \logit\lambda_2 &= \logit\lambda_1 + \epsilon_1 \\\
  \epsilon_1 &\dist N(\mu_{\epsilon_1}, \sigma_{\epsilon_1}^2) \\
  \logit \lambda_1 &\dist N(\mu_{\lambda_1}, \sigma_{\lambda_1}^2) \\
  \sigma &\dist \Exponential(1/\mu_\sigma).
\end{align}
The hyperparmeters $\mu_{\lambda_1}$ and $\sigma_{\lambda_1}$ specify the prior on the hazard at time 1; $\mu_{\epsilon_1}$ and $\sigma_{\epsilon_1}$ the prior on the initial gradient; and $\mu_\sigma$ the smoothness of the random walk.
I use $\mu_{\lambda_1} = -17.5$, $\sigma_{\lambda_1} = 6$, $\mu_\epsilon = 1.09$, $\sigma_\epsilon = 0.03$, and $\mu_\epsilon = 0.1$ based on the analysis in \cref{E-ATACCC}.
\todo[inline]{Add RW2 prior predictive}

>>>>

Smoothing priors encode the biological consideration that the hazard should not vary significantly from day-to-day.
It is the Bayesian equivalent of a penalised likelihood, used frequently in the non-parametric frequentist setting~\autocite[e.g.:][]{bacchettiNonparametric}.
Many possible forms of smoothing priors exist, including splines, Gaussian Processes, and random walks.
I use a second-order random walk (RW2) because it is simple and produces plausible prior predictive distributions.

RW2 prior encodes that the hazard should be linearly changing with some random changes at each time step.
Specifically, the difference on the logit scale between $\lambda_t$ to $\lambda_{t+1}$ is the difference between $\lambda_{t-1}$ and $\lambda_t$ plus some Gaussian-distributed noise.
Formally:
\begin{align}
  \logit\lambda_{t+1}
  &= \logit\lambda_t + (\logit\lambda_t - \logit\lambda_{t-1}) + \sigma \epsilon_t &\text{for $t \geq 2$} \\
  &= 2\logit\lambda_t - \logit\lambda_{t-1} + \sigma \epsilon_t \\
  \epsilon_t &\dist N(0, 1) &\text{for $t \geq 2$}  \\
  \logit\lambda_2 &= \logit\lambda_1 + \epsilon_1 \\\
  \epsilon_1 &\dist N(\mu_{\epsilon_1}, \sigma_{\epsilon_1}^2) \\
  \logit \lambda_1 &\dist N(\mu_{\lambda_1}, \sigma_{\lambda_1}^2) \\
  \sigma &\dist \Exponential(1/\mu_\sigma).
\end{align}
The hyperparmeters $\mu_{\lambda_1}$ and $\sigma_{\lambda_1}$ specify the prior on the hazard at time 1; $\mu_{\epsilon_1}$ and $\sigma_{\epsilon_1}$ the prior on the initial gradient; and $\mu_\sigma$ the smoothness of the random walk.
I use $\mu_{\lambda_1} = -17.5$, $\sigma_{\lambda_1} = 6$, $\mu_\epsilon = 1.09$, $\sigma_\epsilon = 0.03$, and $\mu_\epsilon = 0.1$ based on the analysis in \cref{E-ATACCC} (see \cref{perf-test:fig:rw2-prior}).
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/rw2-prior}
  \caption[RW2 prior for the hazard]{The of prior predictive distribution (mean and 95\% central interval) for the RW2 prior specified in the main text. The posterior mean estimate using ATACCC data in \cref{E-ATACCC} are shown for comparison. \label{perf-test:fig:rw2-prior}}
\end{figure}
